---
title             : "Supplementation of Probiotics on Academic Examination Performance and Stress in Healthy Medical Students"
shorttitle        : "Probiotics Effect on Academic Performance and Stress in Medical Students"

author: 
  - name          : "Nidhi Sheshachala"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role: # Contributorship roles (e.g., CRediT, https://credit.niso.org/)
      - "Conceptualization"
      - "Writing - Original Draft Preparation"
      - "Writing - Review & Editing"
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - "Writing - Review & Editing"
      - "Supervision"

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
source("A2-Package loading and assigning variables.R",local=knitr::knit_global())
```

```{r helloworld,include=TRUE}
source("HelloWorld Function.R", local=knitr::knit_global())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r read probioticsdata}

# Load packages
library(tidyverse)
library(readxl)

# Dataset Read-in
read_excel("Probiotics-Academic-Stress.xlsx")

# Naming Dataset
probiotics_data <- read_excel("Probiotics-Academic-Stress.xlsx")

# Data is tidy
# Each column is a single variable
# Each row is a single observation
# Each cell is a single measurement
# Need to remove individuals who dropped out of the dataset
# Required variables: Participant code, Group enrolment, Drop-out, Sex, Age, STAI trait, Pharmacology examination (points, set, pre-examination test), Anxiety at rest and under stress, Salivary cortisol at rest and under stress, Pulse rate at rest and under stress, salivary 5-HT at rest and under stress.
# Extraneous variables: Group follow-up and completed, Faculty, Nationality, BMI, EAT-26, BDI, and PSS-10
```


``` {r intermediate-probiotics-dataset}
# Using dplyr functions - selecting, filtering, and arranging 
# Displaying only the required variables
im_probiotics_dataset <- probiotics_data |>
  select(`Participant code`, `Group (enrolment)`, `Drop-out`, Sex, `Age [years]`, `STAI trait`, `Pharmacology examination points`, `Pharmacology examination set`, `Pharmacology pre-examination test`, `State anxiety "at rest"`, `State anxiety "under stress"`, `Salivary cortisol "at rest" [ng/ml]`, `Salivary cortisol "under stress" [ng/ml]`, `Pulse rate "at rest" [min-1]`, `Pulse rate "under stress" [min-1]`, `Salivary 5-HT "at rest" [ng/ml]`, `Salivary 5-HT "under stress" [ng/ml]`) |>
  filter(`Drop-out`== 0) |> # Will remove the individuals with "1" indicating they dropped out
  arrange(`Group (enrolment)`) |> # Arranging the dataset to show placebo group and then show Sacch. probiotics group
  print(n=60) # Display all 60 observations

# Exporting the created intermediate dataset
# Load readr
library(readr)
write_csv(im_probiotics_dataset, file =  "im-probiotics-stress-dataset.csv")
```

# Introduction

## Hypothesis

### Definitions

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants
**Healthy Medical Students**

## Material
Probiotic used is _Saccharomyces Boulardii_

<!-- Example of HTML-style comment -->

[Visit RCT Study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7284642/)

## Procedure
### Unordered List:
- Measurement 1
- Measurement 2
- Measurement 3

Measuring the salivary cortisol levels ~~in the body~~

### Ordered List:
1. Step 1
2. Step 2
3. Step 3

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.


# Results

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
