---
title             : "Supplementation of Probiotics on Academic Examination Performance and Stress in Healthy Medical Students"
shorttitle        : "Probiotics Effect on Academic Performance and Stress in Medical Students"

author: 
  - name          : "Nidhi Sheshachala"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "University of Chicago"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

library(tidyverse)
library(readxl)
library(knitr)
library(kableExtra)
library(lme4)
library(psych)
library(stats)
library(scales)
library(usethis)
library(ggh4x)
library(devtools)
devtools::install_github("crsh/citr")
```


```{r final-dataset}
# Dataset Prep

# Dataset read-in - install and load required packages (tidyverse and readxl)
read_excel("Probiotics-Academic-Stress.xlsx")

# Name the dataset
probiotics.dataset <- read_excel("Probiotics-Academic-Stress.xlsx")
colnames(probiotics.dataset) <- tolower(colnames(probiotics.dataset)) # Change all column names to lowercase

# Use dplyr functions to further select, filter and arrange dataset
# Display only the variables required
probiotics.stress.data <- probiotics.dataset %>% 
  filter(`drop-out`== 0) %>% 
  rename(
    "group" = "group (enrolment)",
    "age" = "age [years]",
    "final exam score" = "pharmacology examination points",
    "exam set" = "pharmacology examination set",
    "pre-exam score" = "pharmacology pre-examination test",
    "anxiety at rest" = "state anxiety \"at rest\"",
    "anxiety under stress" = "state anxiety \"under stress\"",
    "cortisol at rest" = "salivary cortisol \"at rest\" [ng/ml]",
    "cortisol under stress" = "salivary cortisol \"under stress\" [ng/ml]",
    "pulse rate at rest" = "pulse rate \"at rest\" [min-1]",
    "pulse rate under stress" = "pulse rate \"under stress\" [min-1]") %>% 
  mutate(`group`= recode(`group`, "P" = "placebo", "S" = "saccharomyces boulardii (probiotic)")) %>% 
  select(`participant code`, `group`, sex, `age`, `pss-10`, `stai trait`, `final exam score`, `exam set`, `pre-exam score`, `anxiety at rest`, `anxiety under stress`, `cortisol at rest`, `cortisol under stress`, `pulse rate at rest`, `pulse rate under stress`) %>% 
  arrange(`group`, `participant code`)

probiotics.stress.data
```

# Introduction
Memory processes are affected by an individual’s experiences, and stress has both positive and negative impacts on memory functioning [@Sandi-memoryimpair, ch. 12]. Buchanan et al., [-@Buchanan-impairedmemory] highglighted that each individual has a unique response and coping mechanism to stressful events, contributing to different effects of stress on memory. Medical students face stressful challenges from the start to the end of their medical journey, and this can impair cognitive ability [@Kar-examstress; @Dyrbye-medstudents]. These students undergo high-stake examinations contributing to performance anxiety [@Dyrbye-medstudents] and increased feelings of stress as the examination day occurs [@Kato-probioticbenefits]. While these stressors can impact academic performance, they may have detrimental effects on cognitive functioning, physiological, and psychological well-being of students. Therefore, it is necessary to identify a variety of effective techniques for stress management whether its mechanism works through improving one’s physical health, emotional functioning, or internally through the gut microbiota [@Worthen-stressmanage]. 

The gut-brain-axis (GBA) highlights the presence of bidirectional communication between the enteric nervous system (ENS) and the central nervous system (CNS) [@Kim-improvedprobioticeffects; @Harvard-probiotics]. With the gut having the ability to modulate brain functioning, targeting the gut microbiome to improve cognition and alleviate stress through probiotic supplementation has been investigated in recent years. Probiotics are strains of live microorganisms that can have beneficial effects on the gut microbiota and are “generally recognized as safe” by the Food and Drug Administration [@Kato-probioticbenefits; @Liu-probioticsdef, p.14; @CCEC-probioticbenefits; @Wauters-probioticbacteria]. Bacterial probiotics are more widely utilized and commonly investigated under research conditions. Probiotic bacterial strains - *Bifidobacterium longum* and *Lactobacillus casei* strain Shirota - have demonstrated health-promoting effects associated with cognitive impairment, academic performance, stress, anxiety and depression [@Kim-improvedprobioticeffects; @Kato-probioticbenefits; @Papalini-stressmatters] in animal and human studies. 

Bacterial and yeast probiotics have unique biological properties, making it difficult to generalize the results obtained from bacterial probiotics to yeast probiotics [@Goplan-probioticproperties; @Pais-probioticinfo]. Saccharomyces boulardii is a yeast probiotic treatment used for gastrointestinal disorders [@Roysarkar-yeastprobioticeffect], and its effects on stress, cognition, academic performance, anxiety and depression are yet to be determined. *S. boulardii* is from the same genus as *Saccharomyces cerevisiae*, otherwise known as baker’s yeast [@Pais-probioticinfo]. This paper aimed to analyze the dataset provided by the study conducted by @Karbownik-datasetstudy examining the effects of *S. boulardii* supplementation on academic performance and stress markers in healthy medical students.

## Hypothesis

### Definitions

## Participants
**Healthy Medical Students**

## Material
Probiotic used is _Saccharomyces Boulardii_

<!-- Example of HTML-style comment -->

[Visit RCT Study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7284642/)

## Procedure
### Unordered List:
- Measurement 1
- Measurement 2
- Measurement 3

Cortisol is a steroid hormone found in the blood, urine, and saliva, and is a common biomarker used to measure physiological stress [@Iqbal-cortisolmarker]. During this study salivary cortisol levels were measured through  the collection of non-invasive saliva samples.


### Ordered List:
1. Step 1
2. Step 2
3. Step 3



``` {r plot1-pre-and-exam-scores, fig.cap="Pre-Exam and Exam Scores with Placebo or Probiotic Treatment"}

# Creates plot1 long data
plot1_data <- gather(probiotics.stress.data, key = "exam_score_type", value = "score", `final exam score`, `pre-exam score`)

# Selecting the variables required for plot1
plot1_columns <- plot1_data %>%
  arrange(desc(exam_score_type)) %>%
  mutate(`exam_score_type` = replace(`exam_score_type`, `exam_score_type` == "pre-exam score", "Pre-exam Score")) %>%
  mutate(`exam_score_type` = replace(`exam_score_type`, `exam_score_type` == "final exam score", "Final Exam Score")) %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "Saccharomyces boulardii (Probiotic)")) %>%
  select(`participant code`, `group`, `exam_score_type`, `score`)

# Re-order the exams so that pre-examination score is on the left side of the plot, and final exam score is on the right side
exam_order <- c("Pre-exam Score", "Final Exam Score")

# Making a boxplot displaying the pre-examination and exam scores for the two different groups (placebo or probiotic)
plot1_columns %>%
  ggplot(aes(x = `group`, y = `score`,  
             fill = `group`)) + 
  geom_boxplot() +
  scale_fill_manual(values = c("Placebo" = "#ffdbff", "Saccharomyces boulardii (Probiotic)" = "#afffff")) +
  scale_color_manual(values = c("Placebo" = "#b81376", "Saccharomyces boulardii (Probiotic)" = "#0083a6")) +
  facet_wrap((~factor(exam_score_type, levels = exam_order)), ncol = 2) +
  geom_point(aes(color = `group`), alpha = 0.8, position = position_jitterdodge()) + # Make sure jitter data points are associated with their respective groups - pre-exam and exam
  labs(title = "Pre-exam and Final Exam Scores for the Placebo and Probiotic Groups",
       x = "",
       y = "Score",
       color = "Group",
       fill = "Group") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_blank(),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    legend.title = element_text(face = "bold", size = 10),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    legend.key = element_rect(fill = 'white'))
```


``` {r plot2-cortisol-and-pulserate, fig.cap="Cortisol and Pulse Rate Changes At Rest and Under Stress with Placebo or Probiotic Treatment."}
plot2_longdata <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "cortisol",
         `cortisol at rest`, `cortisol under stress`) %>%
  gather(key = "at rest or under stress", value = "pulse rate",
         `pulse rate at rest`, `pulse rate under stress`)

plot2_columns <- plot2_longdata %>%
  select('participant code', 'group', 'cortisol', 'at rest or under stress', 'pulse rate') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "Saccharomyces boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("pulse rate at rest", "at rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("pulse rate under stress", "under stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))


plot2_columns %>%
  #filter(!(cortisol > 7.5)) %>%
  ggplot(aes(x = `cortisol`, y = `pulse rate`, shape = `at rest or under stress`, color = `at rest or under stress`)) +
  geom_point() +
  scale_color_manual(values=c("#469B79", "#5D33A7")) +
  #coord_cartesian(xlim = c(0,7.5), ylim = c(40,130)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Salivary Cortisol vs. Pulse Rate of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "Salivary Cortisol [ng/mL]",
       y = "Pulse Rate [min-1]",
       shape = "Group",
       color = "Group") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```


``` {r table1-mean-summary, table.cap="Summary of Mean Academic Performance and Stress Variable Changes At Rest or Under Stress with Placebo or Probiotic Treatment"}
# Create dataset to convert into a table
table_columns <- probiotics.stress.data[, c("participant code", "group", "anxiety at rest", "anxiety under stress", "pre-exam score", "final exam score", "cortisol at rest", "cortisol under stress", "pulse rate at rest", "pulse rate under stress")]

summary_table <- table_columns %>%
  group_by(group) %>% # Was unable to bold the "group" column, will need to modify the dataset
  summarize("Mean Anxiety At Rest" = mean(`anxiety at rest`),
            "Mean Anxiety Under Stress" = mean(`anxiety under stress`),
            "Mean Pre-Exam Score" = mean(`pre-exam score`),
            "Mean Final Exam Score" = mean(`final exam score`),
            "Mean Cortisol At Rest" = mean(`cortisol at rest`),
            "Mean Coritsol Under Stress" = mean(`cortisol under stress`),
            "Mean Pulse Rate At Rest" = mean(`pulse rate at rest`),
            "Mean Pulse Rate Under Stress" = mean(`pulse rate under stress`)) %>%
  pivot_longer(cols = -group, names_to = "Variable", values_to = "Mean") %>%
  pivot_wider(names_from = group, values_from = Mean)

colnames(summary_table) <- c("Variable", "Placebo", "Saccharomyces boulardii (Probiotic)")

summary_table %>%
  kbl(caption = "At Rest and Under Stress Mean Values for Academic Performance and Stress Variables with Placebo or Probiotic Treatment", align = "c") %>%
  kable_styling(full_width = FALSE, font_size = 14, latex_options = c("scale_down")) %>% 
  kable_classic(full_width = F, html_font = "Times New Roman") %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(column = 1:3, border_left = TRUE, border_right = TRUE)
```


```{r anxiety-vs-groups}
plot3_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "anxiety",
         `anxiety at rest`, `anxiety under stress`) %>%
  select('participant code', 'group', 'anxiety', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "Saccharomyces boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("anxiety at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("anxiety under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot3_data %>%
  ggplot(aes(x = `at rest or under stress`, y = `anxiety`, shape = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line() +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Anxiety Levels of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "",
       y = "Anxiety Level",
       shape = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```
``` {r plot4-cortisol-vs-groups}

plot4_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "cortisol",
         `cortisol at rest`, `cortisol under stress`) %>%
  select('participant code', 'group', 'cortisol', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "Saccharomyces boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("cortisol at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("cortisol under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot4_data %>%
  filter(!(cortisol > 10)) %>%
  ggplot(aes(x = `at rest or under stress`, y = `cortisol`, shape = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line() +
  coord_cartesian(ylim = c(0,10)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Salivary Cortisol Levels of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "",
       y = "Cortisol Levels [ng/mL]",
       shape = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```
```{r plot5-pulserate-vs-groups}

plot5_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "pulse rate",
         `pulse rate at rest`, `pulse rate under stress`) %>%
  select('participant code', 'group', 'pulse rate', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "Saccharomyces boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("pulse rate at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("pulse rate under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot5_data %>%
  ggplot(aes(x = `at rest or under stress`, y = `pulse rate`, shape = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line() +
  #coord_cartesian(ylim = c(45,135)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Pulse Rate of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "",
       y = "Pulse Rate [min-1]",
       shape = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```

``` {r hypothesis-testing-cortisol-values}

# Hypothesis Testing Analysis

# Creating the log cortisol dataset
probiotics.stress.data$`log cortisol under stress` <- log(probiotics.stress.data$`cortisol under stress`)
probiotics.stress.data$`log cortisol at rest` <- log(probiotics.stress.data$`cortisol at rest`)

## Testing statistical significance between under stress and at rest cortisol level for the placebo group
placebo_us_at <- probiotics.stress.data %>%
  select(`participant code`, `group`, `log cortisol at rest`, `log cortisol under stress`) %>%
  filter(`group` == "placebo") %>%
  mutate(difference = `log cortisol under stress` - `log cortisol at rest`)

ttest_placebo_us_at <- t.test(
  placebo_us_at$difference
)

# The p value of `r ttest_placebo_us_at$p.value` shows that there is a statistically significant difference 
# between the at rest and under stress cortisol levels for the placebo group.

## Testing statistical significance between under stress and at rest cortisol level for the probiotic group
probiotic_us_at <- probiotics.stress.data %>%
  select(`participant code`, `group`, `log cortisol at rest`, `log cortisol under stress`) %>%
  filter(`group` == "saccharomyces boulardii (probiotic)") %>%
  mutate(difference = `log cortisol under stress` - `log cortisol at rest`)

ttest_probiotic_us_at <- t.test(
  probiotic_us_at$difference
)

## Test statistical significance in the difference between at rest and under stress for the placebo and probiotic groups.
placebo_probiotic_cortisol <- probiotics.stress.data %>%
  mutate(difference = `log cortisol under stress` - `log cortisol at rest`) %>%
  select(`participant code`, `group`, `difference`)

ttest_placebo_probiotic_cortisol <- t.test(
  filter(placebo_probiotic_cortisol, `group` == "placebo")$`difference`,
  filter(placebo_probiotic_cortisol, `group` == "saccharomyces boulardii (probiotic)")$`difference`
)
```


``` {r descriptive-analysis-scores}
# Install and load packages
library(tidyverse)
library(lme4)
library(psych)
library(stats)
library(scales)

# Descriptive Analysis
# Looking at pre-exam and final exam scores for placebo and probiotic groups
library(tidyverse)

# Create dataset with selected variables - 
score.da <- probiotics.stress.data %>%
  gather(key = "exam type", value = "scores",
         `pre-exam score`, `final exam score`) %>%
  select(`group`, `exam type`, `scores`)

score.da %>% # Want to display exam types for both placebo and probiotic groups, and show the summarized descriptive analysis
  group_by(`exam type`, `group`) %>% # Unsure how to change the order of exam type in the analysis - want pre-exam scores before exam scores
  summarize(mean = mean(scores),
            median = median(scores),
            sd = sd(scores),
            range = diff(range(scores)))
```


``` {r hypothesis-testing-analysis-scores-ttest}
# Install and load packages
library(tidyverse)
library(lme4)
library(psych)
library(stats)
library(scales)

# Hypothesis Testing Analysis
# Null hypothesis: There is no difference in exam scores between the placebo and probioitic group.
# Alternative hypothesis: The probiotic group increases exam scores compared to the placebo group.

# Create dataset with selected variables - mainly group and exam score
score.hta <- probiotics.stress.data %>%
  select(`participant code`, `group`, `final exam score`)

# Run t-test
ttest2_overall <- t.test(
  filter(score.hta, `group` == "placebo")$`final exam score`,
  filter(score.hta, `group` == "saccharomyces boulardii (probiotic)")$`final exam score`
)
ttest2_overall
# Not enough statistical significance between the two groups - will be accepting the null hypothesis
# p-value = 0.2312 not significant
```

# Results

# In-line references:
# Methods - Count of Placebo and Probiotic participants in the dataset
This study included `r nrow(filter(probiotics.stress.data, group == "placebo"))` healthy medical students in the Placebo group, and `r nrow(filter(probiotics.stress.data, group == "saccharomyces boulardii (probiotic)"))` healthy medical students in the Saccharomyces boulardii (Probiotic) group.

# Results - Reference the p-value for analysis
The salivary cortisol levels "at rest" and "under stress" were measured for the placebo and probiotic groups. The p-value of `r ttest_placebo_probiotic_cortisol$p.value`, showed that the difference in the "at rest" and "under stress" salivary cortisol levels in the placebo and probiotic groups is not statistically significant. Therefore, the null hypothesis cannot be rejected.

# Count the number of participant IDs that are outliers. 
Defining an outlier as 2.5 * inter quartile range, the number of outliers present during the salivary cortisol under stress condition is `r nrow(filter(probiotics.stress.data, probiotics.stress.data[["cortisol under stress"]] > 2.5 * IQR(probiotics.stress.data[["cortisol under stress"]])))`

# Results - Plot Histogram
The salivary cortisol levels data shows a right skew, as exhibited by `r hist(probiotics.stress.data[["cortisol under stress"]], freq = FALSE, main = "Normal Distribution", xlab = "cortisol under stress", ylab = "Density")`. To make it a normal distribution of values before performing the statistical analysis, we need to log the values.


## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
