---
title             : "Supplementation of Probiotics on Academic Examination Performance and Stress in Healthy Medical Students"
shorttitle        : "Probiotics Effect on Academic Performance and Stress in Medical Students"

author: 
  - name          : "Nidhi Sheshachala"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
affiliation:
  - id            : "1"
    institution   : "University of Chicago"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
  Enter author note here.
  
abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  One or two sentences to put the results into a more **general context**.
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : "r-references.bib"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

library(tidyverse)
library(readxl)
library(knitr)
library(kableExtra)
library(lme4)
library(psych)
library(stats)
library(scales)
library(usethis)
library(ggh4x)
library(devtools)
devtools::install_github("crsh/citr")
```

## Introduction
Memory processes are affected by an individual’s experiences, and stress has both positive and negative impacts on memory functioning [@Sandi-memoryimpair, ch. 12]. Buchanan et al., [-@Buchanan-impairedmemory] highglighted that each individual has a unique response and coping mechanism to stressful events, contributing to different effects of stress on memory. Medical students face stressful challenges from the start to the end of their medical journey, and this can impair cognitive ability [@Kar-examstress; @Dyrbye-medstudents]. These students undergo high-stake examinations contributing to performance anxiety [@Dyrbye-medstudents] and increased feelings of stress as the examination day occurs [@Kato-probioticbenefits]. While these stressors can impact academic performance, they may have detrimental effects on cognitive functioning, physiological, and psychological well-being of students. Therefore, it is necessary to identify a variety of effective techniques for stress management whether its mechanism works through improving one’s physical health, emotional functioning, or internally through the gut microbiota [@Worthen-stressmanage]. 

The gut-brain-axis (GBA) highlights the presence of bidirectional communication between the enteric nervous system (ENS) and the central nervous system (CNS) [@Kim-improvedprobioticeffects; @Harvard-probiotics]. With the gut having the ability to modulate brain functioning, targeting the gut microbiome to improve cognition and alleviate stress through probiotic supplementation has been investigated in recent years. Probiotics are strains of live microorganisms that can have beneficial effects on the gut microbiota and are “generally recognized as safe” by the Food and Drug Administration [@Kato-probioticbenefits; @Liu-probioticsdef, p.14; @CCEC-probioticbenefits; @Wauters-probioticbacteria]. Bacterial probiotics are more widely utilized and commonly investigated under research conditions. Probiotic bacterial strains - *Bifidobacterium longum* and *Lactobacillus casei* strain Shirota - have demonstrated health-promoting effects associated with cognitive impairment, academic performance, stress, anxiety and depression [@Kim-improvedprobioticeffects; @Kato-probioticbenefits; @Papalini-stressmatters] in animal and human studies. 

Bacterial and yeast probiotics have unique biological properties, making it difficult to generalize the results obtained from bacterial probiotics to yeast probiotics [@Goplan-probioticproperties; @Pais-probioticinfo]. Saccharomyces boulardii is a yeast probiotic treatment used for gastrointestinal disorders [@Roysarkar-yeastprobioticeffect], and its effects on stress, cognition, academic performance, anxiety and depression are yet to be determined. *S. boulardii* is from the same genus as *Saccharomyces cerevisiae*, otherwise known as baker’s yeast [@Pais-probioticinfo]. This paper aimed to analyze the dataset provided by the study conducted by @Karbownik-datasetstudy examining the effects of *S. boulardii* supplementation on academic performance and stress markers in healthy medical students.

``` {r dataset-readin}
# Dataset Prep and read-in - load and name the dataset
probiotics.dataset <- read_excel("Probiotics-Academic-Stress.xlsx")
colnames(probiotics.dataset) <- tolower(colnames(probiotics.dataset)) # Change all column names to lowercase
```

```{r final-dataset}
# Use dplyr functions to further select, filter and arrange dataset
# Display only the variables required
probiotics.stress.data <- probiotics.dataset %>% 
  filter(`drop-out`== 0) %>%
  rename(
    "group" = "group (enrolment)",
    "age" = "age [years]",
    "final exam score" = "pharmacology examination points",
    "exam set" = "pharmacology examination set",
    "pre-exam score" = "pharmacology pre-examination test",
    "anxiety at rest" = "state anxiety \"at rest\"",
    "anxiety under stress" = "state anxiety \"under stress\"",
    "cortisol at rest" = "salivary cortisol \"at rest\" [ng/ml]",
    "cortisol under stress" = "salivary cortisol \"under stress\" [ng/ml]",
    "pulse rate at rest" = "pulse rate \"at rest\" [min-1]",
    "pulse rate under stress" = "pulse rate \"under stress\" [min-1]") %>% 
  mutate(`group`= recode(`group`, "P" = "Placebo", "S" = "S. boulardii (probiotic)")) %>% 
  filter(`cortisol under stress` < 80 ) %>%
  select(`participant code`, `group`, sex, `age`, `pss-10`, `stai trait`, `final exam score`, `exam set`, `pre-exam score`, `anxiety at rest`, `anxiety under stress`, `cortisol at rest`, `cortisol under stress`, `pulse rate at rest`, `pulse rate under stress`) %>% 
  arrange(`group`, `participant code`)

probiotics.stress.data
```

## Hypothesis
Null Hypothesis: There is no difference betweenn placebo and *S. boulardii* supplementation on academic performance and stress marker levels in healthy medical students.

Alternative Hypothesis: Supplementation with *S.boulardii* will improve academic performance and decrease stress marker levels in healthy medical students.

## Definitions

## Participants
**Healthy Medical Students**

## Methods
The randomized, double-blind, and placebo-controlled study conducted by @Karbownik-datasetstudy examined the effects of the yeast probiotic *S. boulardii* on academic performance and stress marker levels. Third-year healthy medical students (ages 18-30) in the Faculty of Medicine or Faculty of Military Medicine at the Medical University of Lodz in Poland were recruited as participants for the study. All participation was voluntary.

Participants were randomly assigned the *S. boulardii* (probiotic) or placebo group. This study included `r nrow(filter(probiotics.stress.data, group == "placebo"))` healthy medical students in the Placebo group, and `r nrow(filter(probiotics.stress.data, group == "S. boulardii (probiotic)"))` healthy medical students in the Saccharomyces boulardii (Probiotic) group. Probiotic administration utilized LacidEnter capsules, which contain lyphophilized *S. boulardii* CNCM I-1079, with 5 x 10^9 CFU per dose, while the placebo capusel contained a maltodextrin mixture. 

During this study, the model of stress used was in the form of an academic pharmacology examination.

Cortisol is a steroid hormone found in the blood, urine, and saliva, and is a common biomarker used to measure physiological stress [@Iqbal-cortisolmarker]. During this study salivary cortisol levels were measured through  the collection of non-invasive saliva samples.


[Visit RCT Study](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7284642/)


# Results

``` {r plot1-pre-and-exam-scores, fig.cap="Pre-Exam and Exam Scores with Placebo or Probiotic Treatment"}

# Creates plot1 long data
plot1_data <- gather(probiotics.stress.data, key = "exam_score_type", value = "score", `final exam score`, `pre-exam score`)

# Selecting the variables required for plot1
plot1_columns <- plot1_data %>%
  arrange(desc(exam_score_type)) %>%
  mutate(`exam_score_type` = replace(`exam_score_type`, `exam_score_type` == "pre-exam score", "Pre-exam Score")) %>%
  mutate(`exam_score_type` = replace(`exam_score_type`, `exam_score_type` == "final exam score", "Final Exam Score")) %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  select(`participant code`, `group`, `exam_score_type`, `score`)

# Re-order the exams so that pre-examination score is on the left side of the plot, and final exam score is on the right side
exam_order <- c("Pre-exam Score", "Final Exam Score")

# Making a boxplot displaying the pre-examination and exam scores for the two different groups (placebo or probiotic)
plot1_columns %>%
  ggplot(aes(x = `group`, y = `score`,  
             fill = `group`)) + 
  geom_boxplot() +
  scale_fill_manual(values = c("Placebo" = "#ffdbff", "S. boulardii (Probiotic)" = "#afffff")) +
  scale_color_manual(values = c("Placebo" = "#b81376", "S. boulardii (Probiotic)" = "#0083a6")) +
  facet_wrap((~factor(exam_score_type, levels = exam_order)), ncol = 2) +
  geom_point(aes(color = `group`), alpha = 0.8, position = position_jitterdodge()) + # Make sure jitter data points are associated with their respective groups - pre-exam and exam
  labs(title = "Pre-exam and Final Exam Scores for the Placebo and Probiotic Groups",
       x = "",
       y = "Score",
       color = "Group",
       fill = "Group") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_blank(),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    legend.title = element_text(face = "bold", size = 10),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    legend.key = element_rect(fill = 'white'))
```

As seen in Figure\ \@ref(fig:plot1-pre-and-exam-scores) the pre-exam and exam scores for individuals consuming placebo or probiotics.


``` {r table1-mean-summary, table.cap="Summary of Mean Academic Performance and Stress Variable Changes At Rest or Under Stress with Placebo or Probiotic Treatment"}
# Create dataset to convert into a table
table_columns <- probiotics.stress.data[, c("participant code", "group", "anxiety at rest", "anxiety under stress", "pre-exam score", "final exam score", "cortisol at rest", "cortisol under stress", "pulse rate at rest", "pulse rate under stress")]

summary_table <- table_columns %>%
  group_by(group) %>% # Was unable to bold the "group" column, will need to modify the dataset
  summarize("Mean Anxiety At Rest" = mean(`anxiety at rest`),
            "Mean Anxiety Under Stress" = mean(`anxiety under stress`),
            "Mean Pre-Exam Score" = mean(`pre-exam score`),
            "Mean Final Exam Score" = mean(`final exam score`),
            "Mean Cortisol At Rest" = mean(`cortisol at rest`),
            "Mean Coritsol Under Stress" = mean(`cortisol under stress`),
            "Mean Pulse Rate At Rest" = mean(`pulse rate at rest`),
            "Mean Pulse Rate Under Stress" = mean(`pulse rate under stress`)) %>%
  pivot_longer(cols = -group, names_to = "Variable", values_to = "Mean") %>%
  pivot_wider(names_from = group, values_from = Mean) %>%
  mutate(across(-Variable, ~round(., 1)))

colnames(summary_table) <- c("Variable", "Placebo", "S. boulardii (Probiotic)")

summary_table %>%
  kbl(caption = "At Rest and Under Stress Mean Values for Academic Performance and Stress Variables with Placebo or Probiotic Treatment", align = "c") %>%
  kable_styling(full_width = FALSE, font_size = 14, latex_options = c("scale_down")) %>% 
  kable_classic(full_width = F, html_font = "Times New Roman") %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(column = 1:3, border_left = TRUE, border_right = TRUE)
```

As seen in Table\ \@ref(tab:table1-mean-summary) the summary mean values of academic performance and stress variables for both placebo and probiotic treatment groups.


``` {r plot2-cortisol-and-pulserate, fig.cap="Salivary Cortisol and Pulse Rate Changes At Rest and Under Stress with Placebo or Probiotic Treatment."}

plot2_longdata <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "cortisol",
         `cortisol at rest`, `cortisol under stress`) %>%
  gather(key = "at rest or under stress", value = "pulse rate",
         `pulse rate at rest`, `pulse rate under stress`)

plot2_columns <- plot2_longdata %>%
  select('participant code', 'group', 'cortisol', 'at rest or under stress', 'pulse rate') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("pulse rate at rest", "At rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("pulse rate under stress", "Under stress", `at rest or under stress`)) %>%
  filter(!(cortisol > 15))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot2_columns %>%
  ggplot(aes(x = `cortisol`, y = `pulse rate`, shape = `at rest or under stress`, color = `at rest or under stress`)) +
  geom_point() +
  scale_color_manual(values=c("#00b816", "#007512")) +
  coord_cartesian(xlim = c(0,15), ylim = c(40,130)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Salivary Cortisol vs. Pulse Rate of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "Salivary Cortisol [ng/mL]",
       y = "Pulse Rate [min-1]",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```

As seen in Figure\ \@ref(fig:plot2-cortisol-and-pulserate) the cortisol levels and pulse rates at rest and under stress for the placebo or probiotic groups.


```{r plot3-anxiety-vs-groups, fig.cap="Anxeity Level Changes At Rest and Under Stress with Placebo or Probiotic Treatment."}
plot3_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "anxiety",
         `anxiety at rest`, `anxiety under stress`) %>%
  select('participant code', 'group', 'anxiety', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("anxiety at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("anxiety under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot3_data %>%
  ggplot(aes(x = `at rest or under stress`, y = `anxiety`, shape = `at rest or under stress`, color = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line(size=0.4, colour = "gray") +
  scale_color_manual(values=c("#00b816", "#007512")) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Anxiety Levels of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "",
       y = "Anxiety Level",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```
As seen in Figure\ \@ref(fig:plot3-anxiety-vs-groups) the anxiety levels at rest and under stress for the placebo or probiotic groups.

``` {r plot3-statistical-testing}

## Hypothesis Testing Analysis (Placebo anxiety difference)

# Null hypothesis: There is no difference in anxiety between at rest or under stress.
# Alternative hypothesis: There is a difference in anxiety between at rest or under stress.

# Create dataset with selected variables - group, anxiety at rest and under stress
plot3_stat_data_placebo <- plot3_data %>%
  filter(`group`== "Placebo")
  
# Run t-test
plot3_placebo_ttest <- t.test(
  filter(plot3_stat_data_placebo, `at rest or under stress` == "At Rest")$`anxiety`,
  filter(plot3_stat_data_placebo, `at rest or under stress` == "Under Stress")$`anxiety`
)

# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Probiotic anxiety difference)

# Null hypothesis: There is no difference in anxiety between at rest or under stress.
# Alternative hypothesis: There is a difference in anxiety between at rest or under stress.

# Create dataset with selected variables - group, anxiety at rest and under stress
plot3_stat_data_probiotic <- plot3_data %>%
  filter(`group`== "S. boulardii (Probiotic)")
  
# Run t-test
plot3_probiotic_ttest <- t.test(
  filter(plot3_stat_data_probiotic, `at rest or under stress` == "At Rest")$`anxiety`,
  filter(plot3_stat_data_probiotic, `at rest or under stress` == "Under Stress")$`anxiety`
)

# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Comparing anxiety difference between both groups)

# Null hypothesis: There is no difference in anxiety between both groups.
# Alternative hypothesis: There is a difference in anxiety between both groups.

# Create dataset with selected variables - group, anxiety at rest and under stress
plot3_stat_data_compare <- probiotics.stress.data %>%
  mutate(`anxiety difference` = `anxiety under stress` - `anxiety at rest`) %>%
  select(`participant code`, `group`, `anxiety difference`)
  
# Run t-test
plot3_compare_ttest <- t.test(
  filter(plot3_stat_data_compare, `group` == "Placebo")$`anxiety difference`,
  filter(plot3_stat_data_compare, `group` == "S. boulardii (probiotic)")$`anxiety difference`
)

# No statistical significant difference in anxiety across both groups - will be accepting the null hypothesis
# t score falls within confidence interval
```

Using a Welch Two Sample t-test to study any statistical significance in the change in anxiety levels between at rest and under stress for each group (placebo and probiotic) separately, we see a p-value of `r plot3_placebo_ttest$p.value` and `r plot3_probiotic_ttest$p.value` respectively, portraying a statistically significant difference. Furthermore, investigating the change in difference of anxiety levels between both groups shows no statistically significant result with a p-value of `r plot3_compare_ttest$p.value`.


``` {r plot4-cortisol-vs-groups, fig.cap="Salivary Cortisol Level Changes At Rest and Under Stress with Placebo or Probiotic Treatment."}

plot4_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "cortisol",
         `cortisol at rest`, `cortisol under stress`) %>%
  select('participant code', 'group', 'cortisol', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("cortisol at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("cortisol under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot4_data %>%
  filter(!(cortisol > 15)) %>%
  ggplot(aes(x = `at rest or under stress`, y = `cortisol`, shape = `at rest or under stress`, color = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line(size=0.4, colour = "gray") +
  scale_color_manual(values=c("#00b816", "#007512")) +
  coord_cartesian(ylim = c(0,15)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Salivary Cortisol Levels of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "",
       y = "Cortisol Levels [ng/mL]",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```


```{r plot5-pulserate-vs-groups, fig.cap="Pulse Rate Changes At Rest and Under Stress with Placebo or Probiotic Treatment."}

plot5_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "pulse rate",
         `pulse rate at rest`, `pulse rate under stress`) %>%
  select('participant code', 'group', 'pulse rate', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("pulse rate at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("pulse rate under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot5_data %>%
  ggplot(aes(x = `at rest or under stress`, y = `pulse rate`, shape = `at rest or under stress`, color = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line(size=0.4, colour = "gray") +
  scale_color_manual(values=c("#00b816", "#007512")) +
  #coord_cartesian(ylim = c(45,135)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Pulse Rate of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and Probiotic Groups",
       x = "",
       y = "Pulse Rate [min-1]",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```
As seen in Figure\ \@ref(fig:plot5-pulserate-vs-groups) the anxiety levels at rest and under stress for the placebo or probiotic groups.

``` {r plot5-statistical-testing}

## Hypothesis Testing Analysis (Placebo pulse rate difference)

# Null hypothesis: There is no difference in pulse rate between at rest or under stress.
# Alternative hypothesis: There is a difference in pulse rate between at rest or under stress.

# Create dataset with selected variables - group, pulse rates at rest and under stress
plot5_stat_data_placebo <- plot5_data %>%
  select(`participant code`, `group`, `at rest or under stress`, `pulse rate`) %>%
  filter(`group`== "Placebo")
  
# Run t-test
plot5_placebo_ttest <- t.test(
  filter(plot5_stat_data_placebo, `at rest or under stress` == "At Rest")$`pulse rate`,
  filter(plot5_stat_data_placebo, `at rest or under stress` == "Under Stress")$`pulse rate`
)

# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Probiotic pulse rate difference)

# Null hypothesis: There is no difference in pulse rate between at rest or under stress.
# Alternative hypothesis: There is a difference in pulse rate between at rest or under stress.

# Create dataset with selected variables - group, pulse rates at rest and under stress
plot5_stat_data_probiotic <- plot5_data %>%
  select(`participant code`, `group`, `at rest or under stress`, `pulse rate`) %>%
  filter(`group`== "S. boulardii (probiotic)")
  
# Run t-test
plot5_probiotic_ttest <- t.test(
  filter(plot5_stat_data_probiotic, `at rest or under stress` == "At Rest")$`pulse rate`,
  filter(plot5_stat_data_probiotic, `at rest or under stress` == "Under Stress")$`pulse rate`
)
# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Comparing pulse rate difference between both groups)

# Null hypothesis: There is no difference in pulse rate between both groups.
# Alternative hypothesis: There is a difference in pulse rate between both groups.

# Create dataset with selected variables - group, pulse rates at rest and under stress
plot5_stat_data_compare <- probiotics.stress.data %>%
  mutate(`pulse rate difference` = `pulse rate under stress` - `pulse rate at rest`) %>%
  select(`participant code`, `group`, `pulse rate difference`)
  
# Run t-test
plot5_compare_ttest <- t.test(
  filter(plot5_stat_data_compare, `group` == "Placebo")$`pulse rate difference`,
  filter(plot5_stat_data_compare, `group` == "S. boulardii (probiotic)")$`pulse rate difference`
)
# Statistical significant difference in pulse rate difference across both groups - will be rejecting the null hypothesis
# t score falls outside confidence interval
```

Using a Welch Two Sample t-test to study any statistical significance in the change in pulse rate between at rest and under stress for each group (placebo and probiotic) separately, we see a p-value of `r plot5_placebo_ttest$p.value` and `r plot5_probiotic_ttest$p.value` respectively, portraying a statistically significant difference. Furthermore, investigating the change in increase of pulse rate between both groups shows a statistically significant result with a p-value of `r plot5_compare_ttest$p.value`.


## Data analysis

# Discussion

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::
