---
title             : "Supplementation with Yeast Probiotic, *Saccharomyces Boulardii*, and its Effects on Academic Examination Performance and Stress Markers in Healthy Medical Students"
shorttitle        : "Effects of *S. boulardii* Probiotic on Academic Performance and Stress Levels in Medical Students"

author: 
  - name          : "Nidhi Sheshachala"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    email         : "nidhisb@uchicago.edu"
affiliation:
  - id            : "1"
    institution   : "The University of Chicago"

authornote: |
  The University of Chicago MAPS30550.
  
abstract: Recent research has shed light on the gut microbiome’s impact on brain function through the gut-brain axis (GBA), which is necessary to regulate various host physiological processes. Stressful experiences can disrupt gut microbiome functioning, leading to changes in brain responses and cognitive function. Therefore, it is necessary to restore this gut microbiome regulation during stress to promote cognitive health. Probiotics contain strains of live microorganisms, that can potentially offer beneficial effects to promote overall well-being. While bacterial probiotics like Lactobacillus and Bifidobacterium have shown significant effects in enhancing cognitive function and reducing stress markers, research on yeast probiotics remains limited. This paper aimed to analyze the effects of Saccharomyces boulardii supplementation on academic performance and stress levels in healthy medical students. The participants were randomly assigned either the placebo or S. boulardii (probiotic) group, and assessed their academic performance and stress markers (state anxiety, salivary cortisol, and pulse rate) during “at rest” or “under stress” conditions. While the results did not demonstrate any significant effect on examination performance or stress levels with the administration of  S. boulardii, there was an increase in pulse rate with S. boulardii supplementation, potentially suggesting the body’s body’s “fight or flight” stress response was activated. 
  
keywords          : "Saccharomyces boulardii, yeast probiotic, academic performance, stress, state axiety, salivary cortisol, pulse rate, cognitive function, medical students, gut-brain axis"
wordcount         : "3570"

bibliography      : "r-references.bib"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no

classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

options(repos = c(CRAN = "https://cran.rstudio.com/"))

library(Require)
library(tidyverse)
library(readxl)
library(knitr)
library(kableExtra)
library(lme4)
library(psych)
library(stats)
library(scales)
library(usethis)
Require("ggh4x")
library(devtools)
library(broom)
devtools::install_github("crsh/citr")
```

# Introduction
Memory processes are significantly shaped by an individual’s experiences, and stress can have both positive and negative effects on cognitive functioning [@Sandi-memoryimpair, ch. 12]. Buchanan et al., [-@Buchanan-impairedmemory] demonstrated that each individual responds uniquely to stressful situations, and an inability to effectively cope with excessive stress can detrimentally affect cognition and mental well-being [@AlRouq-medstress; @Sperling-stresshealth]. While stress is a universal experience that affects everyone, medical students represent a demographic particularly susceptible to high levels of stress throughout their academic journey [@Balaji-stressanxiety; @Sperling-stresshealth]. These students face intense pressure associated with high-stakes examinations, leading to performance anxiety [@Dyrbye-medstudentstress]. These stressors can significantly impact the physical and psychological well-being of students, ultimately impairing academic performance and cognitive abilities [@Sperling-stresshealth]. Anxiety and stress often work together to promote cognitive decline, diminishing concentration and working memory among medical students [@Kar-stresseval; @Tianciquek-anxietymedstudent; @Balaji-stressanxiety]. Therefore, it is necessary to explore various effective stress management techniques, whether they operate through enhancing physical health and emotional well-being or internally through modulation of the gut microbiome [@Worthen-stressmanage]. 

The gut microbiome encompasses a diverse community of beneficial commensals - bacteria, fungi, viruses, and other eukaryotes - that work to regulate host homeostatic mechanisms including constant information relay to the brain [@Morkl-probioticsgba]. This bidirectional communication between the enteric nervous system (ENS) and the central nervous system (CNS) termed the gut-brain axis (GBA) [@Kim-probioticimproves; @Morkl-probioticsgba] operates through the activation of mechanisms involving neural, immune, and endocrine pathways [@Takada-probioticlacto]. Particularly during periods of stress, the gut microbiome can impact brain functioning by modulating stress cortisol responses or affecting the hypothalamus-pituitary-adrenal (HPA) axis [@Konturek-stressgut]. However, with exposure to chronic stress, the physiological stress response by the gut microbiome to the HPA axis can be disrupted, impacting cognitive processes in the brain [@AlRouq-medstress]. Given the gut’s ability to influence brain function, there has been a recent focus on investigating probiotic supplementation as a means to enhance cognition and alleviate stress.

Probiotics are strains of live microorganisms dosed in ‘colony forming units’ (CFU)  [@Morkl-probioticsgba] that can have beneficial effects on the gut microbiota and have been recognized as safe by the Food and Drug Administration (FDA) [@Katokataoka-lactoprobiotic; @Liu-probioticsdepression; @Wauters-otherprobioticresults]. Bacterial probiotics, notably *Lactobacillus* and *Bifidobacterium* species are widely utilized and investigated under research conditions. These strains have been associated with improved CNS functioning, during states of gut dysbiosis [@Roysarkar-probioticyeast]. Additionally, *Lactobacillus casei* strain Shirota (LcS) exhibited the ability to reduce salivary cortisol levels and improve physical gastrointestinal (GI) symptoms associated with academic stress [@Takada-probioticlacto]. Similarly, *Bifidobacterium longum* 1714 demonstrated the ability to enhance learning and memory in mouse models exposed to high levels of stress [@Papalini-stressprobiotic]. While the administration of bacterial probiotics has demonstrated potential health-promoting effects on cognitive function, academic performance, stress, and mental well-being [@Kim-probioticimproves; @Dobielska-cogimpaired] in animal and human studies, the effects of yeast probiotics on these aspects remain poorly understood.

The unique properties of probiotics, which vary depending on the species, strain, and variant, make it challenging to apply findings from bacterial probiotics to yeast probiotics [@Goplan-probioticproperties; @Pais-yeastprobiotic]. One such yeast probiotic, *Saccharomyces boulardii*, is commonly used to treat gastrointestinal (GI) disorders [@Roysarkar-probioticyeast; @Pais-yeastprobiotic]. However, its potential effects on cognitive function and stress remain uncertain. *S. boulardii* belongs to the same genus as *Saccharomyces cerevisiae*, known as baker’s yeast [@Pais-yeastprobiotic]. Despite their close relation, *S. boulardii* possesses distinct biological characteristics that make it more advantageous for human health compared to *S. cerevisiae* [@Pais-yeastprobiotic]. For instance, *S. boulardii* thrives at 37°C and exhibits greater resistance to high temperatures, *S. cerevisiae* has an optimal growth temperature of 30°C, which is lower than human host temperatures [@Pais-yeastprobiotic]. Probiotics pass through the GI tract and must survive the adverse conditions posed by gastric juices and bile salts. While bacterial probiotics - *Lactobacillus* and *Bifidobacterium* species - have exhibited decreased viability with exposure to pepsin, hydrochloric acid, and bile salts in the GI tract [@Pais-yeastprobiotic], *S. boulardii* demonstrated the ability to survive the harsh GI conditions. Furthermore, as a yeast-origin probiotic, *S. boulardii* does not pose the risk of genetic transfer, a concern associated with bacterial probiotics [@Goplan-probioticproperties]. Understanding the potential cognitive and stress-related benefits of  *S. boulardii* beyond its GI applications is crucial. To delve into this, the present paper aims to analyze the data from a study conducted by @Karbownik-datasetstudy, examining the effects of *S. boulardii* supplementation on academic performance and stress markers in healthy medical students.

``` {r dataset-readin}
# Dataset Prep and read-in - load and name the dataset
probiotics.dataset <- read_excel("Probiotics-Academic-Stress.xlsx")
colnames(probiotics.dataset) <- tolower(colnames(probiotics.dataset)) # Change all column names to lowercase
```

```{r final-dataset}
# Use dplyr functions to further select, filter and arrange dataset
# Display only the variables required
probiotics.stress.data <- probiotics.dataset %>% 
  filter(`drop-out`== 0) %>%
  rename(
    "group" = "group (enrolment)",
    "age" = "age [years]",
    "final exam score" = "pharmacology examination points",
    "exam set" = "pharmacology examination set",
    "pre-exam score" = "pharmacology pre-examination test",
    "anxiety at rest" = "state anxiety \"at rest\"",
    "anxiety under stress" = "state anxiety \"under stress\"",
    "cortisol at rest" = "salivary cortisol \"at rest\" [ng/ml]",
    "cortisol under stress" = "salivary cortisol \"under stress\" [ng/ml]",
    "pulse rate at rest" = "pulse rate \"at rest\" [min-1]",
    "pulse rate under stress" = "pulse rate \"under stress\" [min-1]") %>% 
  mutate(`group`= recode(`group`, "P" = "Placebo", "S" = "S. boulardii (probiotic)")) %>% 
  filter(`cortisol under stress` < 80 ) %>%
  select(`participant code`, `group`, sex, `age`, `pss-10`, `stai trait`, `final exam score`, `exam set`, `pre-exam score`, `anxiety at rest`, `anxiety under stress`, `cortisol at rest`, `cortisol under stress`, `pulse rate at rest`, `pulse rate under stress`) %>% 
  arrange(`group`, `participant code`)
```

# Methods
The randomized, double-blind, and placebo-controlled study conducted by @Karbownik-datasetstudy examined the effects of the yeast probiotic *S. boulardii* on academic performance and stress marker levels. Third-year healthy medical students (ages 18-30) in the Faculty of Medicine or Faculty of Military Medicine at the Medical University of Lodz in Poland were recruited as participants for the study. All participation was voluntary.

Participants were randomly assigned the *S. boulardii* (probiotic) or placebo group. The study included `r nrow(filter(probiotics.stress.data, group == "Placebo"))` participants in the Placebo group, and `r nrow(filter(probiotics.stress.data, group == "S. boulardii (probiotic)"))` participants in the *S. boulardii* (Probiotic) group. Probiotic administration utilized LacidEnter capsules, which contained lyophilized *S. boulardii* CNCM I-1079, while the placebo capsule contained a maltodextrin mixture (“sugar pill”). Participants in both groups were instructed to consume one capsule daily in the morning for 30 days.

During the study, the model of stress was established in the form of an academic pharmacology exam. The primary outcome evaluated academic performance on the pharmacology final exam, and the secondary outcomes examined the psychology, physiology, and biochemical stress markers. The primary focus measured the number of correctly answered multiple-choice questions in a standard medical pharmacology exam, with a comparison made against a pre-exam assessment completed by participants a day before the final exam. Secondary measures included stress indicators like state anxiety, salivary cortisol levels, and pulse rate. These stress markers were assessed during both pre-intervention no stress (“at rest”) and before the final exam (“under stress”) periods. To evaluate momentary state anxiety levels, the State-Trait Anxiety Inventory (STAI) questionnaire was employed, commonly used as an indicator of distress in stressful conditions research [@Wauters-otherprobioticresults]. Cortisol, a steroid hormone present in the blood, urine, and saliva, served as a prevalent physiological stress marker [@Iqbal-cortisolmarker]. Salivary cortisol levels were measured by collecting non-invasive saliva samples at home.

## Data Analysis
The dataset obtained from the @Karbownik-datasetstudy was filtered to remove any participants who had dropped out of the intervention, as this would not provide complete data to study the statistical significance of pre- and post- *S. boulardii* (probiotic) supplementation. Additionally, only select academic performance and stress marker variables were selected and presented in the format of a long table. The dataset included `r nrow(filter(probiotics.stress.data, group == "Placebo"))` participants in the Placebo group, and `r nrow(filter(probiotics.stress.data, group == "S. boulardii (probiotic)"))` participants in the *S. boulardii* (probiotic) group. 

``` {r print-dataset}
probiotics.stress.data
```

The filtered dataset exhibited two groups - the placebo and *S. boulardii* (probiotic) - which were independent and did not influence each other’s observations. Participants were randomly assigned to either the placebo or *S. boulardii* (probiotic) group and similar sample sizes contributed to the homogeneity of variance of observations. All observations except the salivary cortisol levels displayed approximately normally distributed underlying data. The salivary cortisol variable observations demonstrated a right-skewed distribution due to the influence of outliers present within the dataset. A normal mean salivary cortisol concentration range in the morning is between 3.6 nmol/L (1.44 ng/mL) to 8.3 nmol/L (3.32 ng/mL) [@Iqbal-cortisolmarker]. However, the “cortisol under stress” observations exhibited one participant with a salivary cortisol level of 88.17 ng/mL, which drastically surpasses the mean morning salivary cortisol range. Therefore, to prevent skewing of salivary cortisol results and influencing the statistical significance of any findings, the outlier was removed from the placebo group. Additionally, to further normalize the distribution of salivary cortisol observations, the salivary cortisol [ng/mL] levels were log-transformed. With the previously stated criteria met, statistical hypothesis testing was conducted using Welch’s two-sample t-test. The statistical analysis was conducted at a 95% confidence level, where a p-value < 0.05 was statistically significant.
The correlation coefficient was calculated before conducting further statistical analysis on the slope of the linear regression to identify a statistically significant relationship between salivary cortisol levels and pulse rate.

## Hypothesis
**Null Hypothesis:** There is no difference between placebo and *S. boulardii* supplementation on academic performance and stress levels in healthy medical students.
**Alternative Hypothesis:** Supplementation with *S.boulardii* will enhance academic performance and alter stress marker levels in healthy medical students.
**Note - The *S.boulardii* (Probiotic) group will be referred to as "probiotic" in some sections of the paper.**

# Results

``` {r table1-mean-summary, caption="Summary of Mean Academic Performance and Stress Variable Changes At Rest or Under Stress with Placebo or S. boulardii (Probiotic) Administration.", fig.pos='H', fig.height=3, out.width="100%"}
# Create dataset to convert into a table
table_columns <- probiotics.stress.data[, c("participant code", "group", "anxiety at rest", "anxiety under stress", "pre-exam score", "final exam score", "cortisol at rest", "cortisol under stress", "pulse rate at rest", "pulse rate under stress")]

summary_table <- table_columns %>%
  group_by(group) %>%
  summarize("Mean Anxiety At Rest" = mean(`anxiety at rest`),
            "Mean Anxiety Under Stress" = mean(`anxiety under stress`),
            "Mean Pre-Exam Score" = mean(`pre-exam score`),
            "Mean Final Exam Score" = mean(`final exam score`),
            "Mean Cortisol At Rest" = mean(`cortisol at rest`),
            "Mean Cortisol Under Stress" = mean(`cortisol under stress`),
            "Mean Pulse Rate At Rest" = mean(`pulse rate at rest`),
            "Mean Pulse Rate Under Stress" = mean(`pulse rate under stress`)) %>%
  pivot_longer(cols = -group, names_to = "Variable", values_to = "Mean") %>%
  pivot_wider(names_from = group, values_from = Mean) %>%
  mutate(across(-Variable, ~round(., 1)))

colnames(summary_table) <- c("Variable", "Placebo", "S. boulardii (Probiotic)")

summary_table %>%
  kbl(caption = "Summary of Mean Values for At Rest and Under Stress Academic Performance and Stress Markers with Placebo or S. boulardii (Probiotic) Supplementation", align = "c") %>%
  kable_styling(full_width = FALSE, font_size = 14, latex_options = c("scale_down")) %>% 
  kable_classic(full_width = F, html_font = "Times New Roman") %>%
  row_spec(0, bold = TRUE) %>%
  column_spec(column = 1:3, border_left = TRUE, border_right = TRUE)
```

Table\ \@ref(tab:table1-mean-summary) shows the summary of mean values for academic performance and stress markers during "at rest" and "under stress" for the placebo and *S. boulardii* group.

``` {r plot1-pre-and-exam-scores, fig.cap="Pre-Exam and Final Exam Scores with Placebo or S. boulardii (Probiotic) Supplementation for 30 days.", fig.pos='H', fig.height=4, out.width="100%"}

# Creates plot1 long data
plot1_data <- gather(probiotics.stress.data, key = "exam_score_type", value = "score", `final exam score`, `pre-exam score`)

# Selecting the variables required for plot1
plot1_columns <- plot1_data %>%
  arrange(desc(exam_score_type)) %>%
  mutate(`exam_score_type` = replace(`exam_score_type`, `exam_score_type` == "pre-exam score", "Pre-exam Score")) %>%
  mutate(`exam_score_type` = replace(`exam_score_type`, `exam_score_type` == "final exam score", "Final Exam Score")) %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  select(`participant code`, `group`, `exam_score_type`, `score`)

# Re-order the exams so that pre-examination score is on the left side of the plot, and final exam score is on the right side
exam_order <- c("Pre-exam Score", "Final Exam Score")

# Making a boxplot displaying the pre-examination and exam scores for the two different groups (placebo or probiotic)
plot1_columns %>%
  ggplot(aes(x = `group`, y = `score`,  
             fill = `group`)) + 
  geom_boxplot() +
  scale_fill_manual(values = c("Placebo" = "#ffdbff", "S. boulardii (Probiotic)" = "#afffff")) +
  scale_color_manual(values = c("Placebo" = "#b81376", "S. boulardii (Probiotic)" = "#0083a6")) +
  facet_wrap((~factor(exam_score_type, levels = exam_order)), ncol = 2) +
  geom_point(aes(color = `group`), alpha = 0.8, position = position_jitterdodge()) + # Make sure jitter data points are associated with their respective groups - pre-exam and exam
  labs(title = "Pre-exam and Final Exam Score for Placebo and S. boulardii (Probiotic) Groups",
       x = "",
       y = "Score",
       color = "Group",
       fill = "Group") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 10),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    axis.text.x = element_blank(),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    legend.title = element_text(face = "bold", size = 10),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    legend.key = element_rect(fill = 'white'))
```

``` {r plot1-statistical-testing}

## Hypothesis Testing Analysis (Pre-exam score difference between Placebo and Probiotic groups)

# Create dataset with selected variables - group, pre-exam scores for placebo and probiotic groups.
plot1_preexam_stats <- plot1_data %>%
  select(`participant code`, `group`, `exam_score_type`, `score`) %>%
  filter(`exam_score_type`== "pre-exam score")
  
# Run t-test
plot1_preexam_ttest <- t.test(
  filter(plot1_preexam_stats, `group` == "Placebo")$`score`,
  filter(plot1_preexam_stats, `group` == "S. boulardii (probiotic)")$`score`
)
# No statistical significance between the placebo and probiotic groups - will be accepting the null hypothesis
# t score falls within confidence interval

# Create dataset with selected variables - group, exam scores for placebo and probiotic groups.
plot1_finalexam_stats <- plot1_data %>%
  select(`participant code`, `group`, `exam_score_type`, `score`) %>%
  filter(`exam_score_type`== "final exam score")
  
# Run t-test
plot1_finalexam_ttest <- t.test(
  filter(plot1_finalexam_stats, `group` == "Placebo")$`score`,
  filter(plot1_finalexam_stats, `group` == "S. boulardii (probiotic)")$`score`
)
# No statistical significance between the placebo and probiotic groups - will be accepting the null hypothesis
# t score falls within confidence interval
```

The pre-examination and final examination scores of participants in the placebo and probiotic groups was displayed in Figure\ \@ref(fig:plot1-pre-and-exam-scores). The pre-exam scores between the placebo and probiotic groups were compared and exhibited a p-value of `r plot1_preexam_ttest$p.value`, indicating no statistical significance. Comparing the final exam scores for both groups also displayed no statistical significance, with a p-value of `r plot1_finalexam_ttest$p.value`. This is further evident when investigating the confidence interval of `r plot1_preexam_ttest$conf.int` for the pre-exam scores and `r plot1_finalexam_ttest$conf.int` since 0.00 falls within these ranges. Therefore, *S. boulardii* supplementation did not significantly enhance final exam scores compared to the placebo group, leading us to accept the null hypothesis.

``` {r plot2-cortisol-and-pulserate, fig.cap="Salivary Cortisol [ng/mL] vs. Pulse Rate [min-1] Changes At Rest and Under Stress with Placebo or S. boulardii (Probiotic) Administration.", fig.pos='H', fig.height=4, out.width="\\textwidth"}

plot2_longdata <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "cortisol",
         `cortisol at rest`, `cortisol under stress`) %>%
  gather(key = "at rest or under stress", value = "pulse rate",
         `pulse rate at rest`, `pulse rate under stress`)

plot2_columns <- plot2_longdata %>%
  select('participant code', 'group', 'cortisol', 'at rest or under stress', 'pulse rate') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("pulse rate at rest", "At rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("pulse rate under stress", "Under stress", `at rest or under stress`)) %>%
  filter(!(cortisol > 15))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot2_columns %>%
  ggplot(aes(x = `cortisol`, y = `pulse rate`, shape = `at rest or under stress`, color = `at rest or under stress`)) +
  geom_point() +
  scale_color_manual(values=c("#00b816", "#007512")) +
  coord_cartesian(xlim = c(0,15), ylim = c(40,130)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Salivary Cortisol vs. Pulse Rate Changes in Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and S. boulardii (Probiotic) Groups",
       x = "Salivary Cortisol [ng/mL]",
       y = "Pulse Rate [min-1]",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    panel.border = element_rect(color = "black", fill = NA),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```

```{r plot2-statistical-testing}

## Placebo at rest
plot2_placebo_data_atrest <- plot2_columns %>%
  filter(group =="Placebo", `at rest or under stress` == "At rest")

plot2_placebo_lm_atrest <- tidy(lm(`pulse rate` ~ `cortisol`, data = plot2_placebo_data_atrest))

# p value for slope is 0.13895. No statistically significant evidence between cortisol and pulse rate.

## Placebo under stress
plot2_placebo_data_understress <- plot2_columns %>%
  filter(group =="Placebo", `at rest or under stress` == "Under stress")

plot2_placebo_lm_understress <- tidy(lm(`pulse rate` ~ `cortisol`, data = plot2_placebo_data_understress))

# p value for slope is 0.1252089. No statistically significant evidence between cortisol and pulse rate.


## Probiotic at rest
plot2_probiotic_data_atrest <- plot2_columns %>%
  filter(group =="S. boulardii (Probiotic)", `at rest or under stress` == "At rest")

plot2_probiotic_lm_atrest <- tidy(lm(`pulse rate` ~ `cortisol`, data = plot2_probiotic_data_atrest))

# p value for slope is 0.9202859. No statistically significant evidence between cortisol and pulse rate.

## Probiotic under stress
plot2_probiotic_data_understress <- plot2_columns %>%
  filter(group =="S. boulardii (Probiotic)", `at rest or under stress` == "Under stress")

plot2_probiotic_lm_understress <- tidy(lm(`pulse rate` ~ `cortisol`, data = plot2_probiotic_data_understress))

# p value for slope is 0.7684893. No statistically significant evidence between cortisol and pulse rate.
```

```{r plot2_correlation}
# Correlation coefficient between Cortisol and Pulse rate - at rest and placebo group
cor_atrest_placebo <- plot2_columns %>%
  filter(`group` == "Placebo", `at rest or under stress` == "At rest") %>%
  select(`cortisol`, `pulse rate`) %>%
  cor()

# Correlation coefficient between Cortisol and Pulse rate - at rest and probiotic group
cor_atrest_probiotic <- plot2_columns %>%
  filter(`group` == "S. boulardii (Probiotic)", `at rest or under stress` == "At rest") %>%
  select(`cortisol`, `pulse rate`) %>%
  cor()

# Correlation coefficient between Cortisol and Pulse rate - under stress and placebo group
cor_understress_placebo <- plot2_columns %>%
  filter(`group` == "Placebo", `at rest or under stress` == "Under stress") %>%
  select(`cortisol`, `pulse rate`) %>%
  cor()

# Correlation coefficient between Cortisol and Pulse rate - under stress and probiotic group
cor_understress_probiotic <- plot2_columns %>%
  filter(`group` == "S. boulardii (Probiotic)", `at rest or under stress` == "Under stress") %>%
  select(`cortisol`, `pulse rate`) %>%
  cor()
```

Figure\ \@ref(fig:plot2-cortisol-and-pulserate) examined whether there was a correlation between salivary cortisol levels [ng/mL] and pulse rate [min-1] during "at rest" and "under stress" conditions for both the placebo and *S. boulardii* (probiotic) group. For the placebo group, the correlation coefficient between cortisol and pulse rate "at rest" was `r cor_atrest_placebo[2]` and `r cor_understress_placebo[2]` "under stress". For the probiotic group, the correlation coefficient was `r cor_atrest_probiotic[2]` "at rest" and `r cor_understress_probiotic[2]` "under stress".
The slope for "at rest" and "under stress" conditions was calculated for each treatment group. The placebo slope values were not statistically significant, with an "at rest" slope value of `r plot2_placebo_lm_atrest$p.value[2]`, and an "under stress" slope value of `r plot2_placebo_lm_understress$p.value[2]`. Similarly the *S. boulardii* (probiotic) group also displayed no statistical significance, with a slope value of `r plot2_probiotic_lm_atrest$p.value[2]` "at rest", and slope value of `r plot2_probiotic_lm_understress$p.value[2]` "under stress." As both groups did not display any statistical significance, and given the low correlation coefficients, there is no strong evidence to show a correlation between salivary cortisol levels and pulse rate changes in this healthy medical student sample.

```{r plot3-anxiety-vs-groups, fig.cap="State Anxeity Changes At Rest and Under Stress with Placebo or S. boulardii (Probiotic) Supplementation.", fig.pos='H', fig.height=3.5, out.width="100%"}
plot3_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "anxiety",
         `anxiety at rest`, `anxiety under stress`) %>%
  select('participant code', 'group', 'anxiety', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("anxiety at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("anxiety under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot3_data %>%
  ggplot(aes(x = `at rest or under stress`, y = `anxiety`, shape = `at rest or under stress`, color = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line(linewidth=0.4, colour = "gray") +
  scale_color_manual(values=c("#00b816", "#007512")) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "State Anxiety Levels of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and S. boulardii (Probiotic) Groups",
       x = "",
       y = "Anxiety Level",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```

``` {r plot3-statistical-testing}

## Hypothesis Testing Analysis (Placebo anxiety difference)

# Null hypothesis: There is no difference in anxiety between at rest or under stress.
# Alternative hypothesis: There is a difference in anxiety between at rest or under stress.

# Create dataset with selected variables - group, anxiety at rest and under stress
plot3_stat_data_placebo <- plot3_data %>%
  filter(`group`== "Placebo")
  
# Run t-test
plot3_placebo_ttest <- t.test(
  filter(plot3_stat_data_placebo, `at rest or under stress` == "At Rest")$`anxiety`,
  filter(plot3_stat_data_placebo, `at rest or under stress` == "Under Stress")$`anxiety`
)

# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Probiotic anxiety difference)

# Null hypothesis: There is no difference in anxiety between at rest or under stress.
# Alternative hypothesis: There is a difference in anxiety between at rest or under stress.

# Create dataset with selected variables - group, anxiety at rest and under stress
plot3_stat_data_probiotic <- plot3_data %>%
  filter(`group`== "S. boulardii (Probiotic)")
  
# Run t-test
plot3_probiotic_ttest <- t.test(
  filter(plot3_stat_data_probiotic, `at rest or under stress` == "At Rest")$`anxiety`,
  filter(plot3_stat_data_probiotic, `at rest or under stress` == "Under Stress")$`anxiety`
)

# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Comparing anxiety difference between both groups)

# Null hypothesis: There is no difference in anxiety between both groups.
# Alternative hypothesis: There is a difference in anxiety between both groups.

# Create dataset with selected variables - group, anxiety at rest and under stress
plot3_stat_data_compare <- probiotics.stress.data %>%
  mutate(`anxiety difference` = `anxiety under stress` - `anxiety at rest`) %>%
  select(`participant code`, `group`, `anxiety difference`)
  
# Run t-test
plot3_compare_ttest <- t.test(
  filter(plot3_stat_data_compare, `group` == "Placebo")$`anxiety difference`,
  filter(plot3_stat_data_compare, `group` == "S. boulardii (probiotic)")$`anxiety difference`
)

# No statistical significant difference in anxiety across both groups - will be accepting the null hypothesis
# t score falls within confidence interval
```

The state anxiety levels "at rest" and "under stress" for the placebo and *S. boulardii* (probiotic) group is demonstrated in Figure\ \@ref(fig:plot3-anxiety-vs-groups). Statistical significance in the change in state anxiety levels between "at rest" and "under stress" for the placebo and *S. boulardii* (probiotic) group was first conducted separately. Both groups showed a statistically significant difference where the placebo group had a p-value of `r format(plot3_placebo_ttest$p.value, digits = 3)`, and the *S. boulardii* (probiotic) showed a p-value of `r format(plot3_probiotic_ttest$p.value, digits = 3)`. The t-scores portrayed a significant change in the mean anxiety level by `r abs(plot3_placebo_ttest$statistic)` standard deviations in the placebo group and `r abs(plot3_probiotic_ttest$statistic)` in the probiotic group. Furthermore, the confidence intervals are `r abs(plot3_placebo_ttest$conf.int)` for the placebo group and `r abs(plot3_probiotic_ttest$conf.int)` suggesting a statistically significant difference between the "at rest" and "under stress" anxiety levels across both groups.
Furthermore, when investigating the change in difference of state anxiety levels between the two groups, there was no statistically significant result as a p-value of `r format(plot3_compare_ttest$p.value, digits = 2)` is observed. Therefore, although there is a statistically significant increase in anxiety, the change is consistent across both groups with 0.00 falling within the confidence interval of `r plot3_compare_ttest$conf.int`. The null hypothesis is accepted, as there is no difference in state anxiety levels with *S. boulardii* (probiotic) use.

``` {r logplot4-cortisol-vs-groups, fig.cap="Salivary Cortisol Level Changes At Rest and Under Stress with Placebo or S. boulardii (Probiotic) Administration for 30 Days.", fig.pos='H', fig.height=3.5, out.width="100%"}

plot4_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "cortisol",
         `cortisol at rest`, `cortisol under stress`) %>%
  select('participant code', 'group', 'cortisol', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "S. boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("cortisol at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("cortisol under stress", "Under Stress", `at rest or under stress`)) %>%
  mutate(`log cortisol` = log(cortisol))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot4_data %>%
  filter(!(cortisol > 15)) %>%
  ggplot(aes(x = `at rest or under stress`, y = `log cortisol`, shape = `at rest or under stress`, color = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line(linewidth=0.4, colour = "gray") +
  scale_color_manual(values=c("#00b816", "#007512")) +
  coord_cartesian(ylim = c(-1,3)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Salivary Cortisol Levels of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and S. boulardii (Probiotic) Groups",
       x = "",
       y = "Log Cortisol Levels log[ng/mL]",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```

``` {r plot4-statistical-testing}

## Hypothesis Testing Analysis (Placebo cortisol difference)

# Null hypothesis: There is no difference in cortisol between at rest or under stress.
# Alternative hypothesis: There is a difference in cortisol between at rest or under stress.

# Create dataset with selected variables - group, cortisol at rest and under stress
plot4_stat_data_placebo <- plot4_data %>%
  filter(`group`== "Placebo")
  
# Run t-test
plot4_placebo_ttest <- t.test(
  filter(plot4_stat_data_placebo, `at rest or under stress` == "At Rest")$`log cortisol`,
  filter(plot4_stat_data_placebo, `at rest or under stress` == "Under Stress")$`log cortisol`
)
# No statistical significance between the at rest and under stress - will be accepting the null hypothesis
# p value < 0.05 and 0 within confidence interval

## Hypothesis Testing Analysis (Probiotic cortisol difference)

# Null hypothesis: There is no difference in cortisol between at rest or under stress.
# Alternative hypothesis: There is a difference in cortisol between at rest or under stress.

# Create dataset with selected variables - group, cortisol at rest and under stress
plot4_stat_data_probiotic <- plot4_data %>%
  filter(`group`== "S. boulardii (Probiotic)")
  
# Run t-test
plot4_probiotic_ttest <- t.test(
  filter(plot4_stat_data_probiotic, `at rest or under stress` == "At Rest")$`log cortisol`,
  filter(plot4_stat_data_probiotic, `at rest or under stress` == "Under Stress")$`log cortisol`
)
# No statistical significance between the at rest and under stress - will be accepting the null hypothesis
# p value < 0.05 and 0 within confidence interval

## Hypothesis Testing Analysis (Comparing cortisol difference between both groups)

# Null hypothesis: There is no difference in cortisol between both groups.
# Alternative hypothesis: There is a difference in cortisol between both groups.

# Create dataset with selected variables - group, cortisol at rest and under stress
plot4_stat_data_compare <- probiotics.stress.data %>%
  mutate(`cortisol difference` = `cortisol under stress` - `cortisol at rest`) %>%
  select(`participant code`, `group`, `cortisol difference`)
  
# Run t-test
plot4_compare_ttest <- t.test(
  filter(plot4_stat_data_compare, `group` == "Placebo")$`cortisol difference`,
  filter(plot4_stat_data_compare, `group` == "S. boulardii (probiotic)")$`cortisol difference`
)
# No statistical significant difference in cortisol across both groups - will be accepting the null hypothesis
# p value < 0.05 and 0 within confidence interval
```

"At rest" and "under stress" changes in salivary cortisol levels [ng/mL] in placebo and *S. boulardii* (probiotic) groups were observed in Figure\ \@ref(fig:logplot4-cortisol-vs-groups). Creating a histogram to show the distribution of the underlying salivary cortisol levels across both groups in the "at rest" and "under stress" conditions depicted a heavy right skew. This prompted a log-transformation to derive a near-normal distribution for further statistical testing. Null hypothesis assumed no difference in cortisol levels between the "at rest" or "under stress" states and the alternative hypothesis a two-tailed test for a potentially higher or lower cortisol levels "under stress". Given that the data was log-transformed, the geometric means were compared in the analysis. The analysis on the placebo group presented a t-score of `r plot4_placebo_ttest$statistic`, suggesting a decrease in geometric mean cortisol level of `r plot4_placebo_ttest$statistic` standard deviations "under stress" compared to the "at rest" state. However, this was not a statistically significant result with a p-value of `r plot4_placebo_ttest$p.value` and with 0.00 falling within the confidence interval of `r plot4_placebo_ttest$conf.int`. A similar finding was seen with the *S. boulardii* (probiotic) group with a t-score of `r plot4_probiotic_ttest$statistic`, a p-value of `r plot4_probiotic_ttest$p.value`, and 0.00 falling within the confidence interval of `r plot4_probiotic_ttest$conf.int`.
To analyze the statistical significance in the cortisol levels between the placebo and probiotic groups, cortisol "at rest" was subtracted from cortisol "under stress" for each participant. No log-transformation was attempted due to a few negative results which cannot be computed mathematically. This showed that although the mean cortisol was `r plot4_compare_ttest$statistic` standard deviations higher in the probiotic group, the p-value of `r plot4_compare_ttest$p.value` and 0 within the confidence interval of `r plot4_compare_ttest$conf.int` reiterates the statistical insignificance in the finding, reaffirming the acceptance of the null hypothesis.

```{r plot5-pulserate-vs-groups, fig.cap="Pulse Rate Changes At Rest and Under Stress with Placebo or S. boulardii (Probiotic) Administration.", fig.pos='H', fig.height=3.5, out.width="100%"}

plot5_data <- probiotics.stress.data %>%
  gather(key = "at rest or under stress", value = "pulse rate",
         `pulse rate at rest`, `pulse rate under stress`) %>%
  select('participant code', 'group', 'pulse rate', 'at rest or under stress') %>%
  mutate(`group` = replace(`group`, `group` == "placebo", "Placebo")) %>%
  mutate(`group` = replace(`group`, `group` == "saccharomyces boulardii (probiotic)", "S. boulardii (Probiotic)")) %>%
  mutate(`at rest or under stress` = gsub("pulse rate at rest", "At Rest", `at rest or under stress`)) %>%
  mutate(`at rest or under stress` = gsub("pulse rate under stress", "Under Stress", `at rest or under stress`))

strip_colors <-strip_themed(background_x = elem_list_rect(fill = c("#ffdbff", "#afffff")))

plot5_data %>%
  ggplot(aes(x = `at rest or under stress`, y = `pulse rate`, shape = `at rest or under stress`, color = `at rest or under stress`, group = `participant code`)) +
  geom_point() +
  geom_line(linewidth=0.4, colour = "gray") +
  scale_color_manual(values=c("#00b816", "#007512")) +
  #coord_cartesian(ylim = c(45,135)) +
  facet_wrap2(~ `group`, scales = "free", strip = strip_colors) +
  labs(title = "Pulse Rate of Healthy Medical Students",
       subtitle = "At Rest and Under Stress Changes in the Placebo and S. boulardii (Probiotic) Groups",
       x = "",
       y = "Pulse Rate [min-1]",
       shape = "Condition",
       color = "Condition") +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    plot.subtitle = element_text(size = 10),
    axis.title = element_text(size = 10),
    legend.background = element_rect(color = "black", fill = "white"),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    strip.text = element_text(face = "bold", size = 8, color = "black"),
    legend.key = element_rect(fill = 'white'),
    axis.text.x = element_text(size = 8),
    axis.text.y = element_text(size = 8))
```

``` {r plot5-statistical-testing}

## Hypothesis Testing Analysis (Placebo pulse rate difference)

# Null hypothesis: There is no difference in pulse rate between at rest or under stress.
# Alternative hypothesis: There is a difference in pulse rate between at rest or under stress.

# Create dataset with selected variables - group, pulse rates at rest and under stress
plot5_stat_data_placebo <- plot5_data %>%
  select(`participant code`, `group`, `at rest or under stress`, `pulse rate`) %>%
  filter(`group`== "Placebo")
  
# Run t-test
plot5_placebo_ttest <- t.test(
  filter(plot5_stat_data_placebo, `at rest or under stress` == "At Rest")$`pulse rate`,
  filter(plot5_stat_data_placebo, `at rest or under stress` == "Under Stress")$`pulse rate`
)
# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Probiotic pulse rate difference)

# Null hypothesis: There is no difference in pulse rate between at rest or under stress.
# Alternative hypothesis: There is a difference in pulse rate between at rest or under stress.

# Create dataset with selected variables - group, pulse rates at rest and under stress
plot5_stat_data_probiotic <- plot5_data %>%
  select(`participant code`, `group`, `at rest or under stress`, `pulse rate`) %>%
  filter(`group`== "S. boulardii (probiotic)")
  
# Run t-test
plot5_probiotic_ttest <- t.test(
  filter(plot5_stat_data_probiotic, `at rest or under stress` == "At Rest")$`pulse rate`,
  filter(plot5_stat_data_probiotic, `at rest or under stress` == "Under Stress")$`pulse rate`
)
# Statistical significance between the at rest and under stress - will be rejecting the null hypothesis
# t score falls outside confidence interval

## Hypothesis Testing Analysis (Comparing pulse rate difference between both groups)

# Null hypothesis: There is no difference in pulse rate between both groups.
# Alternative hypothesis: There is a difference in pulse rate between both groups.

# Create dataset with selected variables - group, pulse rates at rest and under stress
plot5_stat_data_compare <- probiotics.stress.data %>%
  mutate(`pulse rate difference` = `pulse rate under stress` - `pulse rate at rest`) %>%
  select(`participant code`, `group`, `pulse rate difference`)
  
# Run t-test
plot5_compare_ttest <- t.test(
  filter(plot5_stat_data_compare, `group` == "Placebo")$`pulse rate difference`,
  filter(plot5_stat_data_compare, `group` == "S. boulardii (probiotic)")$`pulse rate difference`
)
# Statistical significant difference in pulse rate difference across both groups - will be rejecting the null hypothesis
# t score falls outside confidence interval
```

Pulse rate [min-1] changes "at rest" and "under stress" in the placebo and *S. boulardii* (probiotic) groups were illustrated in Figure\ \@ref(fig:plot5-pulserate-vs-groups). The placebo group displayed a p-value of `r format(plot5_placebo_ttest$p.value, digits = 2)`, indicating a statistically significant change between "at rest" and "under stress" pulse rate levels. A statistically significant p-value of `r format(plot5_probiotic_ttest$p.value, scientific = FALSE, digits = 2)` was also observed in the *S. boulardii* (probiotic) group. This is reiterated by the confidence interval of `r abs(plot5_placebo_ttest$conf.int)` for the placebo group and `r abs(plot5_probiotic_ttest$conf.int)` for the probiotic group.
When comparing the placebo and *S. boulardii* (probiotic) groups, a statistically significant result (p-value = `r format(plot5_compare_ttest$p.value, digits = 2)`) in pulse rate changes can allow the null hypothesis to be rejected.

# Discussion
Bacterial probiotics have emerged as potential candidates for mitigating stress-related markers and enhancing cognitive functions through the modulation of the GBA. However, the exploration of similar effects with yeast probiotics is limited in the scientific literature [@Goplan-probioticproperties]. The current paper analyzed the data obtained from the @Karbownik-datasetstudy study, which investigated the impact of supplementation with the yeast probiotic *Saccharomyces boulardii* on academic performance and stress markers among healthy medical students. The analysis revealed that the administration of  *S. boulardii* (probiotic) for 30 days did not display significant improvements in academic performance or a decrease in stress-related markers under exam-induced stress conditions.
Given the short-termed duration of *S. boulardii’s* presence in the gut, typically lasting no more than 3-5 days post-ingestion, and its documented effectiveness in addressing GI disorders [@Pais-yeastprobiotic], it was proposed that it might similarly enhance academic performance and alleviate stress, as observed with bacterial probiotics [@Karbownik-datasetstudy]. While *S. boulardii* demonstrates the ability to survive the harsh conditions posed by the GI tract, yeast organisms are unable to strongly colonize the colon, resulting in its fast clearance from the GI tract in healthy individuals [@Pais-yeastprobiotic]. As this intervention was conducted on healthy medical students, a potential reason for the absence of significant effects on academic performance and stress with *S. boulardii* use could be due to its limited ability to effectively compete with established commensal microbiota in the gut and its restricted adherence to intestinal epithelial cells [@Pais-yeastprobiotic]. 
Since different probiotic strains possess distinct characteristics, their efficacy may vary across different demographic groups [@Morkl-probioticsgba]. Although supplementation of *S. boulardii* did not result in significant effects on academic performance and stress levels among healthy medical students, it's plausible that alternative effects might manifest in different demographic cohorts. Notably, @Pais-yeastprobiotic reported that the concentration of *S. boulardii* increased in the gut following antibiotic administration, suggesting that *S. boulardii’s* effects may be more pronounced under conditions of gut microbiome depletion. Furthermore, individual responses to *S. boulardii* could be influenced by variations in the unique gut microbiome composition of each individual, which can be shaped by genetic, dietary, and other demographic factors (ethnicity, health status etc.) [@Dobielska-cogimpaired]. Despite the exclusion of individuals following a vegan or vegetarian diet during participant selection, the study did not report participants’ daily dietary intake. Consumption of fermented products consisting of *Lactobacillus* and *Bifidobacterium* species can alter the microbial composition and potentially inhibit the effects of *S. boulardii* [@Pais-yeastprobiotic]. Therefore, considering the significant impact of diet on gut microbiome composition, future investigations should monitor participant diets to examine whether dietary factors influence the effects of *S. boulardii* in improving cognitive functioning and reducing stress levels.
@Takada-probioticlacto indicated that GI or flu-like symptoms often accompany stressful situations, and the administration of LcS (a bacterial probiotic) demonstrated the ability to alleviate these physical symptoms during academic examination stress conditions. However, the study conducted by @Karbownik-datasetstudy did not report the presence of GI or cold-like symptoms associated with an academic examination (stressor). Given that *S. boulardii* is used for the treatment of GI disorders [@Roysarkar-probioticyeast], there is potential for this yeast probiotic to alleviate GI symptoms linked to exam-related stress. Consequently, future investigations should monitor changes in GI symptoms during “at rest” and “under stress” conditions following *S. boulardii* supplementation.
While the overall findings of the study indicated that *S. boulardii* did not have any significant effects on the primary and secondary outcomes, Figure\ \@ref(fig:plot5-pulserate-vs-groups) surprisingly revealed a statistically significant increase in pulse rate in both the placebo (p-value = `r format(plot5_placebo_ttest$p.value, digits = 2)`) and probiotic (p-value = `r format(plot5_probiotic_ttest$p.value, scientific = FALSE, digits = 2)`) groups. The rejection of the null hypothesis was based on a statistically significant p-value of `r format(plot5_compare_ttest$p.value, digits = 2)`, obtained from comparing the two groups. In the *S. boulardii* (probiotic) group, the mean pulse rate went from 69.5 [min-1] “at rest” to 91.3 [min-1] “under stress” (Table\ \@ref(tab:table1-mean-summary)). Additionally, during the “under stress” conditions the mean pulse rate for the placebo group was 85.1 [min-1], while supplementation with *S. boulardii* (probiotic) led to an elevated mean pulse rate of 91.3 [min-1]. These results suggest the potential of *S. boulardii* (probiotic) to activate the sympathoadrenal system, initiating a “fight or flight” response [@Karbownik-datasetstudy]. This heightened sympathoadrenal response during exam-related stressors could potentially elevate adrenaline and cortisol levels, which can negatively impact academic performance [@Karbownik-datasetstudy]. Furthermore, with caffeine consumption being prevalent among medical students, diet may have contributed to this observed increased pulse rate, as caffeine can elevate blood pressure and stimulate stress hormone secretion [@Hartley-caffeinestress]. Caffeine may have synergistically increased the effect of *S. boulardii* on pulse rate through various host-regulated mechanisms. Further examination is essential to validate the associated between *S. boulardii* supplementation and increase pulse rate, as well as to elucidate the underlying mechanisms employed by this yeast probiotic. A thorough understanding of the pathways activated by *S. boulardii* to elevate pulse rate could facilitate the assessment of its therapeutic potential in cardiovascular function.
The small sample size could have hindered the detection of statistical significance and limited the generalizability of the findings. Also, the study only examined the effects of *S. boulardii* strain CNCM I-1079, highlighting the necessity for future investigations to explore whether different strains (*S. boulardii* CNCM-1745) [@Karbownik-datasetstudy], elicit a similar impact on academic performance and stress levels.

# Conclusion
Overall, *S. boulardii* supplementation for 30 days did not have a statistically significant effect on enhancing academic performance or reducing stress-related marker levels in healthy medical students, indicating a potential disparity in therapeutic ability compared to bacterial probiotics for cognitive enhancement. Further research exploring different strains, larger sample sizes, and *S. boulardii’s* effects on elevated pulse rate are essential to fully understand *S. boulardii’s* potential therapeutic benefits with cognitive functioning and stress.


\newpage
# References 

::: {#refs custom-style="Bibliography"}
:::
